Looking for a Signal in the Noise: Revisiting Obesity and the microbiome

### 2016-03-24 Sze_ObesityMicrobiomeMetaNoMetaHit_PeerJ_2016.Rmd
another manuscript update

demographics <- read.csv("results/tables/denovodemographicsTable.csv")
OverallPValue <- read.csv("results/tables/denovoOverallPTable.csv")
ShannonRRTable <- read.csv("results/tables/denovoShannonRRTable.csv")
BFRatioRRTable <- read.csv("results/tables/denovoBFRatioRRTable.csv")
ZScoreCombinedData <- read.csv("results/tables/denovoCombinedData.csv")
ZScoreTurnbaughData <- read.csv("results/tables/denovoTurnbaughData.csv")
AUCRFData <- read.csv("results/tables/denovoAUCRFDataTable.csv")
PowerTable <- read.csv("results/tables/denovoPowerTable.csv")
ModelledPower1PerDiff <- read.csv("results/tables/denovoModelled1PerDiff.csv")
ModelledPower5PerDiff <- read.csv("results/tables/denovoModelled5PerDiff.csv")
ModelledPower10PerDiff <- read.csv("results/tables/denovoModelled10PerDiff.csv")
ModelledPower15PerDiff <- read.csv("results/tables/denovoModelled15PerDiff.csv")
ActualDiff <- read.csv("results/tables/denovoActualDiff.csv")
PowerModelHRR <- read.csv("results/tables/denovoPowerModelHRR.csv")
PowerModelBFRR <- read.csv("results/tables/denovoPowerModelBFRR.csv")
SimulatedN_H <- read.csv("results/tables/denovoSimulatedN_H.csv")
SimulatedN_BF <- read.csv("results/tables/denovoSimulatedN_BF.csv")
RRSimNNeeded <- read.csv("results/tables/denovoRRSimNNeeded.csv")

@@

@@

table1 <- cbind(Study, AgeTable, Sex, Ancestry, BMI, Min, Max)
table1[table1=="NA" | table1=="NA | NA" | table1=="NA (NA)" | 
         table1=="      NA"] <- "Not Available"

colnames(table1) <- c("Study", "Age (SD)", "Sex (F|M)", "European Ancestry (%)", "BMI (SD)", "Min BMI","Max BMI")

@@

**Table 3. Summary of Power Calculation Simulations for Shannon Diversity for Non-obese versus Obese**

```{r Table_3, echo=FALSE, message=FALSE, warning=FALSE}
ObeseTotal <- ShannonRRTable$tposH + ShannonRRTable$cposH
ContTotal <- ShannonRRTable$tnegH + ShannonRRTable$cnegH

Study <- paste(as.character(PowerTable$Study), " (", ContTotal, "|", ObeseTotal, ")", sep="")
ActualDiffH <- format(round(ActualDiff$H, 3), nsmall=3)
Shannon <- format(round(OverallPValue$Shannon, 2), nsmall=2)
OneP <- format(round(ModelledPower1PerDiff$H, 3), nsmall=3)
FiveP <- format(round(ModelledPower5PerDiff$H, 3), nsmall=3)
TenP <- format(round(ModelledPower10PerDiff$H, 3), nsmall=3)
FifteenP <- format(round(ModelledPower15PerDiff$H, 3), nsmall=3)

table3 <- cbind(Study, ActualDiffH, Shannon, OneP, FiveP, TenP, FifteenP)
colnames(table3) <- c("Study (Non-Obese|Obese)", "Actual Percent Difference", "P-value", "Estimated Power for 1% Difference", "Estimated Power for 5% Difference", "Estimated Power for 10% Difference", "Estimated Power for 15% Difference")

kable(table3, format = "markdown", align = 'c', padding = 2)
```

**Table 4. Summary of Power Calculation Simulations for B/F Ratio for Non-obese versus Obese**

```{r Table_4, echo=FALSE, message=FALSE, warning=FALSE}

Study <- paste(as.character(PowerTable$Study), " (", ContTotal, "|", ObeseTotal, ")", sep="") 
ActualDiffBF <- format(round(ActualDiff$BFratio, 3), nsmall=3)
BF <- format(round(OverallPValue$BFRatio, 2), nsmall=2)
OneP <- format(round(ModelledPower1PerDiff$BFratio, 3), nsmall=3)
FiveP <- format(round(ModelledPower5PerDiff$BFratio, 3), nsmall=3)
TenP <- format(round(ModelledPower10PerDiff$BFratio, 3), nsmall=3)
FifteenP <- format(round(ModelledPower15PerDiff$BFratio, 3), nsmall=3)

table4 <- cbind(Study, ActualDiffBF, BF, OneP, FiveP, TenP, FifteenP)
colnames(table4) <- c("Study (Non-Obese|Obese)", "Actual Percent Difference", "P-value", "Estimated Power for 1% Difference", "Estimated Power for 5% Difference", "Estimated Power for 10% Difference", "Estimated Power for 15% Difference")

kable(table4, format = "markdown", align = 'c', padding = 2)

**Table 5. Summary of Sample Size Needed Simulations for Shannon Diversity for Non-obese versus Obese**

@@

Study <- paste(as.character(PowerTable$Study), " (", ContTotal, "|", ObeseTotal, ")", sep="")

NNeeded <- format(round(SimulatedN_H[, -1],0), nsmall=0)
table5 <- cbind(Study, NNeeded)
colnames(table5) <- c("Study (Non-Obese|Obese)", "Actual n/Group Needed", "Estimated n/Group for 1% Difference", "Estimated n/Group for 5% Difference", "Estimated n/Group for 10% Difference", "Estimated n/Group for 15% Difference")

kable(table5, format = "markdown", align = 'c', padding = 2)
```

**Table 6. Summary of Sample Size Needed Simulations for B/F Ratio for Non-obese versus Obese**

@@

Study <- paste(as.character(PowerTable$Study), " (", ContTotal, "|", ObeseTotal, ")", sep="")

NNeeded <- format(round(SimulatedN_BF[, -1],0), nsmall=0)
table6 <- cbind(Study, NNeeded)
colnames(table6) <- c("Study (Non-Obese|Obese)", "Actual n/Group Needed", "Estimated n/Group for 1% Difference", "Estimated n/Group for 5% Difference", "Estimated n/Group for 10% Difference", "Estimated n/Group for 15% Difference")

kable(table6, format = "markdown", align = 'c', padding = 2)
```

@@

```{r Figure_4, echo=FALSE, message=FALSE, warning=FALSE, fig.path="results/figures/"}
# Section A
p1 <- ggplot(group.data.test, aes(obese, ZH))
p2 <- p1 + geom_boxplot(aes(fill = obese)) + scale_fill_brewer(palette="Set1")
p3 <- p2 + theme(axis.text=element_text(size=8), axis.title=element_text(size=10, face="bold"))
p4 <- p3 + xlab("") + ylab("Z-score Normalized H") + scale_x_discrete(breaks=c("No", "Yes"), labels=c("Non-Obese", "Obese"))
p5 <- p4 + guides(fill=guide_legend(title="Obese")) + theme(legend.title = element_text(size=8, face="bold"))
p6 <- p5 + theme(axis.line=element_line(colour="black"),
                 panel.grid.major = element_blank(),
                 panel.grid.minor = element_blank(),
                 panel.border = element_blank(),
                 panel.background = element_blank())
p7 <- p6 + ggtitle('A') + theme(plot.title=element_text(hjust=0, face="bold", size=20))

# Section B
q1 <- ggplot(group.data.test, aes(BMI.class, ZH))
q2 <- q1 + geom_boxplot(aes(fill = BMI.class)) + scale_fill_brewer(palette="Set1")
q3 <- q2 + theme(axis.text=element_text(size=8), axis.title=element_text(size=10, face="bold"))
q4 <- q3 + xlab("") + ylab("Z-score Normalized H")
q5 <- q4 + guides(fill=guide_legend(title="BMI Class")) + theme(legend.title = element_text(size=8, face="bold"))
q6 <- q5 + theme(axis.line=element_line(colour="black"),
                 panel.grid.major = element_blank(),
                 panel.grid.minor = element_blank(),
                 panel.border = element_blank(),
                 panel.background = element_blank())
q7 <- q6 + ggtitle('B') + theme(plot.title=element_text(hjust=0, face="bold", size=20))

# Section C
bf1 <- ggplot(group.data.test, aes(obese, ZLogBF))
bf2 <- bf1 + geom_boxplot(aes(fill = obese)) + scale_fill_brewer(palette="Set1")
bf3 <- bf2 + theme(axis.text=element_text(size=8), axis.title=element_text(size=10, face="bold"))
bf4 <- bf3 + xlab("") + ylab("Z-score Normalized Log B/F Ratio") + 
  scale_x_discrete(breaks=c("No", "Yes"), labels=c("Non-Obese", "Obese"))
bf5 <- bf4 + guides(fill=guide_legend(title="Obese")) + theme(legend.title = element_text(size=8, face="bold"))
bf6 <- bf5 + theme(axis.line=element_line(colour="black"),
                   panel.grid.major = element_blank(),
                   panel.grid.minor = element_blank(),
                   panel.border = element_blank(),
                   panel.background = element_blank())
bf7 <- bf6 + ggtitle('C') + theme(plot.title=element_text(hjust=0, face="bold", size=20))

# Section D
qbf1 <- ggplot(group.data.test, aes(BMI.class, ZLogBF))
qbf2 <- qbf1 + geom_boxplot(aes(fill = BMI.class)) + scale_fill_brewer(palette="Set1")
qbf3 <- qbf2 + theme(axis.text=element_text(size=8), axis.title=element_text(size=10, face="bold"))
qbf4 <- qbf3 + xlab("") + ylab("Z-score Normalized Log B/F Ratio")
qbf5 <- qbf4 + guides(fill=guide_legend(title="BMI Class")) + theme(legend.title = element_text(size=8, face="bold"))
qbf6 <- qbf5 + theme(axis.line=element_line(colour="black"),
                     panel.grid.major = element_blank(),
                     panel.grid.minor = element_blank(),
                     panel.border = element_blank(),
                     panel.background = element_blank())
qbf7 <- qbf6 + ggtitle('D') + theme(plot.title=element_text(hjust=0, face="bold", size=20))

grid.arrange(p7, q7, bf7, qbf7)
```

>>>>

demographics <- read.csv("results/tables/denovodemographicsTable.csv")
OverallPValue <- read.csv("results/tables/denovoOverallPTable.csv")
ShannonRRTable <- read.csv("results/tables/denovoShannonRRTable.csv")
BFRatioRRTable <- read.csv("results/tables/denovoBFRatioRRTable.csv")
ZScoreCombinedData <- read.csv("results/tables/denovoCombinedData.csv")
ZScoreTurnbaughData <- read.csv("results/tables/denovoTurnbaughData.csv")
AUCRFData <- read.csv("results/tables/denovoAUCRFDataTable.csv")
PowerTable <- read.csv("results/tables/denovoPowerTable.csv")
ModelledPower1PerDiff <- read.csv("results/tables/denovoModelled1PerDiff.csv")
ModelledPower5PerDiff <- read.csv("results/tables/denovoModelled5PerDiff.csv")
ModelledPower10PerDiff <- read.csv("results/tables/denovoModelled10PerDiff.csv")
ModelledPower15PerDiff <- read.csv("results/tables/denovoModelled15PerDiff.csv")
ActualDiff <- read.csv("results/tables/denovoActualDiff.csv")
PowerModelHRR <- read.csv("results/tables/denovoPowerModelHRR.csv")
PowerModelBFRR <- read.csv("results/tables/denovoPowerModelBFRR.csv")
SimulatedN_H <- read.csv("results/tables/denovoSimulatedN_H.csv")
SimulatedN_BF <- read.csv("results/tables/denovoSimulatedN_BF.csv")
RRSimNNeededH <- read.csv("results/tables/denovoRRSimNNeededH.csv")
RRSimNNeededBF <- read.csv("results/tables/denovoRRSimNNeededBF.csv")

@@

ObBreakdown <- paste(ContTotal, " | ", ObeseTotal, sep="")

@@

table1 <- cbind(Study, AgeTable, Sex, Ancestry, ObBreakdown, BMI, Min, Max)
table1[table1=="NA" | table1=="NA | NA" | table1=="NA (NA)" | 
         table1=="      NA"] <- "Not Available"

colnames(table1) <- c("Study", "Age (SD)", "Sex (F|M)", "European Ancestry (%)", "Obese (No|Yes)", "BMI (SD)", "Min BMI","Max BMI")

@@

@@

source("code/MakeFiguresCode.R")
grid.arrange(a2, b2, nrow=2)

```
**Figure 5. Summary of Power and Sample Size Simulations for Shannon Diversity for Non-obese versus Obese**

@@

grid.arrange(c2, d2, nrow=2)

```
**Figure 6. Summary of Power and Sample Size Simulations for Bacteroidetes/Firmicutes Ratio for Non-obese versus Obese**

@@

```{r Figure_4, echo=FALSE, message=FALSE, warning=FALSE, fig.path="results/figures/"}
# Section A
p1 <- ggplot(group.data.test, aes(obese, ZH))
p2 <- p1 + geom_boxplot(aes(fill = obese)) + scale_fill_brewer(palette="Set1")
p3 <- p2 + theme(axis.text=element_text(size=8), axis.title=element_text(size=10, face="bold"))
p4 <- p3 + xlab("") + ylab("Z-score Normalized H") + scale_x_discrete(breaks=c("No", "Yes"), labels=c("Non-Obese", "Obese"))
p5 <- p4 + guides(fill=FALSE)
p6 <- p5 + theme(axis.line=element_line(colour="black"),
                 panel.grid.major = element_blank(),
                 panel.grid.minor = element_blank(),
                 panel.border = element_blank(),
                 panel.background = element_blank())
p7 <- p6 + ggtitle('A') + theme(plot.title=element_text(hjust=0, face="bold", size=20))

# Section B
bf1 <- ggplot(group.data.test, aes(obese, ZLogBF))
bf2 <- bf1 + geom_boxplot(aes(fill = obese)) + scale_fill_brewer(palette="Set1")
bf3 <- bf2 + theme(axis.text=element_text(size=8), axis.title=element_text(size=10, face="bold"))
bf4 <- bf3 + xlab("") + ylab("Z-score Normalized Log B/F Ratio") + 
  scale_x_discrete(breaks=c("No", "Yes"), labels=c("Non-Obese", "Obese"))
bf5 <- bf4 + guides(fill=FALSE)
bf6 <- bf5 + theme(axis.line=element_line(colour="black"),
                   panel.grid.major = element_blank(),
                   panel.grid.minor = element_blank(),
                   panel.border = element_blank(),
                   panel.background = element_blank())
bf7 <- bf6 + ggtitle('B') + theme(plot.title=element_text(hjust=0, face="bold", size=20))

grid.arrange(p7, bf7, ncol=2)
```
